---
title: "STAT 405 Project"
author: "Group 4: Safa Eltgani, Jumi Lee, Juliet Mankin, Tapasya Nangpal, Breanna Schneider"
output:
  html_document:
    df_print: paged
  pdf_document: default
fontsize: 10pt
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	cache = TRUE
)
```
[Link to Github Repository](https://github.com/bjschneider4/405project)

### Project Proposal  
Using the 31G flight data set from Kaggle, we will analyze ticket prices based on 
time interval between when a flight is searched for and its departure date based on  departing airport. This approach mirrors the "best time to book" recommendations provided by platforms like Google Flights, aiming to offer actionable insights for both consumers and airlines. We hope to provide flyers with the best possible price using information beyond the "best time to book" output. Our analysis will explore the implications of number of seats remaining, origin and landing destinations, as well as day of the week. The data is from Expedia between 2022-04-16 and 2022-10-05. 
The airport arrival/destinations are ATL, DFW, DEN, ORD, LAX, CLT, MIA, JFK, 
EWR, SFO, DTW, BOS, PHL, LGA, IAD, OAK. It has a total of 16 departure airports. 


### Description of Variables : 
`legId`: An identifier for the flight.    
`searchDate`: The date (YYYY-MM-DD) on which this entry was taken from Expedia.   
`flightDate`: The date (YYYY-MM-DD) of the flight.    
`startingAirport`: Three-character IATA airport code for the initial location.    
`destinationAirport`: Three-character IATA airport code for the arrival location.   
`fareBasisCode`: The fare basis code.   
`travelDuration`: The travel duration in hours and minutes.   
`elapsedDays`: The number of elapsed days (usually 0).    
`isBasicEconomy`: Boolean for whether the ticket is for basic economy.    
`isRefundable`: Boolean for whether the ticket is refundable.   
`isNonStop`: Boolean for whether the flight is non-stop.    
`baseFare`: The price of the ticket (in USD).   
`totalFare`: The price of the ticket (in USD) including taxes and other fees.   
`seatsRemaining`: Integer for the number of seats remaining.    
`totalTravelDistance`: The total travel distance in miles. This data is sometimes missing.    
`segmentsDepartureTimeEpochSeconds`: String containing the departure time (Unix time) for each leg of the trip.   
`segmentsDepartureTimeRaw`: String containing the departure time.     
`segmentsArrivalTimeEpochSeconds`: String containing the arrival time (Unix time) for each leg of the trip.
`segmentsArrivalTimeRaw`: String containing the arrival time  for each leg of the trip.    
`segmentsArrivalAirportCode`: String containing the IATA airport code for the arrival location for each leg of the trip.     
`segmentsDepartureAirportCode`: String containing the IATA airport code for the departure location for each leg of the trip.     
`segmentsAirlineName`: String containing the name of the airline that services each leg of the trip.     
`segmentsAirlineCode`: String containing the two-letter airline code that services each leg of the trip.   
`segmentsEquipmentDescription`: String containing the type of airplane used for each leg of the trip.   
`segmentsDurationInSeconds`: String containing the duration of the flight (in seconds) for each leg of the trip.     
`segmentsDistance`: String containing the distance traveled (in miles) for each leg of the trip.    
`segmentsCabinCode`: String containing the cabin for each leg of the trip.     

### Code to Read Data:  
[Link to data](https://www.kaggle.com/datasets/dilwong/flightprices)
```{r,eval=FALSE}

require(gt)
require(data.table)
flights=fread("itineraries.csv",nrows=1000)
head(flights)

flight=flights[,-c(1:3,5:8,12,16:27)]
head(flight,n=2) %>%
  gt()
```


### Statistical Methods
We will use multiple linear regression to predict the price of a flight.These models are ideal for predicting continuous variables—in our case, flight prices—based on multiple independent variables, such as time to departure, time booked, route, and potentially other relevant factors found within the dataset. Our outcome variable will simply be baseFare to avoid variation in state level taxes and varies security fees. To prevent overgeneralizing and provide relevant information to passengers, we will develop a model for each origin airport. Due to the large number of potential explanatory variables, feature selection is going to be an initial challenge. We hope to experiment with each of the above variables (accounting for redundant ones like `segmentsArrivalTimeEpochSeconds` and `segmentsArrivalTimeRaw`) to see which have the most meaningful impact on the model. Afterwards, we will provide a reduced model with the significant features and run it on subsetted data for each origin airport and airline.  

### Computational Steps
1. Split data into several smaller files by departing airport

2. Use best subset or some other variable selection method to chose which variables are significant to price.

3. Fit a model and run diagnostics to evaluate model performance and validity, ensuring the models accurately capture the dynamics of flight pricing.

Steps 2 and 3 will be distributed to the departing airport files.

